A Gentle Introduction into Structural Causal Models

### 2021-04-30 paper/draft.Rmd
yz

    number_sections: yes
    fig_caption: yes
citation_package: --biblatex
fontsize: 12
linestretch: 1.5
header-includes:
- \usepackage{amsmath}
- \usepackage[makeroom]{cancel}
abstract: This paper provides a gentle introduction into causal inference focusing on structural causal models. Structural Causal Models are abstractions of underlying physical mechanisms. SCMs are used for causal inference. This paper briefly introduces the role of time in causality. I will use a two and three variable model to illustrate underlying assumptions of structural causal models.
---

@@

>>>>

    number_sections: yes
    fig_caption: yes
citation_package: --biblatex
fontsize: 12
linestretch: 1.5
header-includes:
- \usepackage{amsmath}
- \usepackage[makeroom]{cancel}
abstract: Interest in a deeper understand of relationships between variables has increased the popularity of causal models. There are various different causal model specifications.
I will explore two models namely (1) Bayesian causal networks (BCN) and (2) structural causal models (SCM). Probabilisitic specifications such as a BCN cast a model based on conditional probabilities. SCMs cast a model based on functions. SCMs extend probabilisitic models by specifying the data generating process rather than solely focusing on conditional probabilities. These SCMs can be applied to number of different queries. Using Pearl's hierachy of causation (2009), I focus on prediction, intervention and counterfactuals. To further clarify central notions of causal modelling, this paper briefly discusses the role of time in causality and will illustrate an general overview of the history of causal modelling to contexualize the SCM modelling approach. I will also introduce causal graphs, specifically focusing on the popular directed acyclical graphical (DAG) models, to highlight differences in various queries and modelling assumptions.
---

@@

"Correlation does not imply Causation" is one of the most pivotal lessons many students learn during introductory undergraduate statistics courses.
